Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-06-04 source/sec_intro.tex
Update on Overleaf.

\subsection{Related Works}
\textbf{The Convergence of FedAvg and Linear Speedup.}
There have been a considerable amount of convergence results in distributed stochastic optimization with infrequent communications that establish linear speedup in the number of clients. However, most of these prior works provide
convergence guarantees under simplified settings, where either statistical
heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} is not
considered, and the associated algorithm is often referred to as local SGD. ~\cite{li2019convergence} is the first to provide convergence of FedAvg for general strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied, whereas this paper focuses on convex settings.
\begin{comment}
However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
FedAvg. 
\end{comment}

@@

% Our contributions
% \subsection{Contributions}
\textbf{Contributions.}
We provide a comprehensive convergence analysis
of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity.
Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
More specifically, for strongly convex and smooth problems, we show 
that FedAvg admits a convergence rate of 
{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
Last but not least, we systematically examine the convergence rate of
FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.

>>>>

\subsection{Related Works}
\textbf{Convergence of FedAvg}
There have been a considerable amount of convergence results in distributed stochastic optimization with infrequent communications that establish linear speedup in the number of clients. However, most of these prior works provide
convergence guarantees under simplified settings, where either statistical
heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} is not
considered, and the associated algorithm is often referred to as local SGD. ~\cite{li2019convergence} is the first to provide convergence of FedAvg for general strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied, whereas this paper focuses on convex settings.
\begin{comment}
However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
FedAvg. 
\end{comment}

@@

% Our contributions
% \subsection{Contributions}
\subsection{Our Contributions.}
We provide a comprehensive convergence analysis
of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity.
Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
More specifically, for strongly convex and smooth problems, we show 
that FedAvg admits a convergence rate of 
{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
Last but not least, we systematically examine the convergence rate of
FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.
